---
title: "QRISK_meta_analysis"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
  html_document:
    df_print: paged
header-includes:
- \usepackage{fvextra}
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,breakanywhere,commandchars=\\\{\}}
---

install.packages("readxl") install.packages("meta") install.packages("metafor") install.packages("dplyr") install.packages("ggplot2")

# IMPORT DATA FROM EXCEL FILE

```{r}
# Inport the data from the xcel fille
# Load package
library(readxl)

# Google Drive file ID (uploaded .xlsx, not a native Sheet)
sheet_id <- "1NjldJRQ_qM6wM4eAfa2cxahBVpHxiB25"
url <- paste0("https://drive.google.com/uc?export=download&id=", sheet_id)

tmp <- tempfile(fileext = ".xlsx")
download.file(url, tmp, mode = "wb")     # mode = "wb" is essential
df <- read_excel(tmp)

# View first few rows
head(df)

# Check structure
str(df)

# Check column names
names(df)
```

CLEAN THE IMPORTED DATA BY REMOVING ANY EMPTY COLUMNS THE RECHECK DATA

```{r}
# Remove columns that are entirely NA
df <- df[, colSums(!is.na(df)) > 0]

# Check dimensions
dim(df)

# Check column names
names(df)
```

REMOVE COMPLETELY EMPTY ROWS

```{r}
# Remove rows where all values are NA
library(dplyr)
df <- df %>%
  filter(!if_all(everything(), is.na))
```

CHECK THE FORMAT OF INPUTED CONFIDENCE INTERVAL IN ORDER TO DEVICE PLAN TO PARSE IT

```{r}
# check CI format
df$AUC[!is.na(df$AUC)][1:20]
```

# SPLIT CONFIDENCE INTERVAL

CLEAN AND EXTRACT NUMERICAL VALUES FROM AUC COLUMN

```{r}
# extract CI into lower and upper column 
library(stringr) # Provides fuction for manipulating text
library(dplyr) # provides function for manipulating data

df <- df %>% # use the pipe operator to pass the result of manipulation and passes it to the right
  mutate( # create new columns while retaining exisiting ones
    AUC_clean = str_squish(AUC), #remove unnecessary spaces from AUC and store it in new variable

    Estimate = str_extract(AUC_clean, "^\\d*\\.?\\d+") |> as.numeric(), # creates a variabl, Estimate and store the first number seen in the AUC_clean is it as a number

    LowerCI = str_match( # Using regular expression extract the lower confidence interval and store it in LowerCI column
      AUC_clean,
      "\\(\\s*(\\d*\\.?\\d+)\\s*[–-]\\s*(\\d*\\.?\\d+)\\s*\\)"
    )[,2] |> as.numeric(),

    UpperCI = str_match( # Using regular expression extract the Upper confidence interval and store it in UpperCI column
      AUC_clean,
      "\\(\\s*(\\d*\\.?\\d+)\\s*[–-]\\s*(\\d*\\.?\\d+)\\s*\\)"
    )[,3] |> as.numeric()
  )
```

```{r}
# view all the new extracted data
df %>%
  select(AUC, Estimate, LowerCI, UpperCI) %>%
  head(30)
```

## HANLEY & MCNEIL (1982) APPROXIMATION

```{r}
# HANLEY & MCNEIL (1982) APPROXIMATION
library(dplyr)

df_hm <- df # create a working copy

df_hm <- df_hm %>%
  mutate( # create a new column
    n1 = as.numeric(Events), # store the information in the Event column in n1 as number
    n0 = Sample_size - n1, # compute the number of none events by subtracting event from sample size
    Q1 = Estimate / (2 - Estimate), # compute for Q1
    Q2 = 2 * Estimate^2 / (1 + Estimate), # compute for Q2
    HM_se    = sqrt((Estimate * (1 - Estimate) + # h and M original variance estimator
                     (n1 - 1) * (Q1 - Estimate^2) +
                     (n0 - 1) * (Q2 - Estimate^2)) / (n1 * n0)),
    HM_lower = pmax(0, Estimate - 1.96 * HM_se), # calcualte lower confidence interval
    HM_upper = pmin(1, Estimate + 1.96 * HM_se) # calculate higher confidence interval
  )

# Output the values that where affected by the approximation
hm_table <- df_hm %>% # what would be
  filter(Confidence_interval == "Approximated") %>% # isolate row that currently do not have reported confidence interval
  transmute(
    Study    = Index, # indentify the row by the index values
    Events   = n1,
    AUC      = round(Estimate, 3),
    `95% CI` = sprintf("%.3f – %.3f", HM_lower, HM_upper)
  )

print(as.data.frame(hm_table), row.names = FALSE)
```

# CALCULATE STANDARD ERROR

```{r}
# STANDARD ERROR
df$SE <- (df$UpperCI - df$LowerCI) / 3.92 
```

```{r}
# VERIFY NO VLAUES ARE MISSING
sum(is.na(df$SE))
summary(df$SE)
```

### SPLIT DATA (df_all, df_primary)

df_all : dataframe containing all included studies

df_primary : dataframe containing only studies included in the primary pool

```{r}
# Load the package
library(writexl)
```

```{r}
# create df with all data including QRISK
df_all <- df

nrow(df_all)
head(df_all)
```

```{r}
# Export the dataframe
write_xlsx(df_all, "df_all.xlsx")
```

Extracting the primary pool studies

```{r}
# create dataframe for primary pooled AUC
# keep only studies marked for inclusion
df_primary <- df[df$Include == 1, ]

# keep only studies marked for meta - analysis
df_primary <- df_primary[df_primary$Metanalysis == 1, ]

# remove hellers k studies
df_primary <- df_primary[df_primary$Dscrimination != "Heller’s K", ]

df_primary <- df_primary[df_primary$Outcome_type == "Appropriate", ]

# View first rows
nrow(df_primary)
str(df_primary$Index)
```

```{r}
write_xlsx(df_primary, "df_primary.xlsx")
```

TRANSFORM EACH C-STATISTIC TO LOGIT SCALE FOR DEBRAY'S ADHERENCE

```{r}
# transform the AUC values from their original probability scale (0–1) to the logit scale and calculates the standard error (SE) on the logit scale.
library(dplyr)

df_primary_logit <- df_primary %>%
  mutate(
    logit_c  = log(Estimate / (1 - Estimate)), # create logit transformed AUC
    logit_lb = log(LowerCI  / (1 - LowerCI)),
    logit_ub = log(UpperCI  / (1 - UpperCI)),
    se_logit = (logit_ub - logit_lb) / (2 * 1.96)   # derive the standard error of the logit c-statistic 
  )
```

```{r}
# Export the dataframe
write_xlsx(df_primary_logit, "df_primary-logit.xlsx")
```

```{r}
# TO CHECK
# output unique studies for primary metanalysis
library(dplyr)
library(stringr)

# Eligible cohorts
eligible <- df_primary_logit %>%
  filter(Metanalysis == 1) %>%
  mutate(
    StudyID = str_extract(as.character(Index), "\\d+")
  )

nrow(eligible)


cat(
  "Number of unique studies:",
  n_distinct(eligible$StudyID),
  "\n"
)

eligible %>%
  select(Index, Title) %>%
  print(n = Inf)
```

Count the number of arms contributed by each study

```{r}
df_primary_logit %>%
  mutate(StudyID = str_extract(as.character(Index), "^[0-9]+")) %>% # CONVERT index to character and then pick out characters from the begining until you find a character that is not a number and store that in studyID
  count(StudyID, name = "n_arms") %>% # get the frequency of each number and store in the column n_arms
  arrange(desc(n_arms)) %>% # arrange the result in decending order
  print(n = Inf) # print all rows
```

# META-ANALYSIS STARTS

```{r}
# LOAD META PACKAGE
library(meta)
```

```{r}
# INVERSE VARIANCE RANDOM EFFECT Using non logit estimates
primary_auc <- metagen( # use metagen package to combine AUC estimates and store in created object
  TE = Estimate, # effect estimate gotten from Estimate column
  seTE = SE, # get the treatment effect standard error from the SE column
  studlab = Author, # use the Author column to identify studies
  data = df_primary, # dataset that contains all the data
  sm = "AUC", # specify the type of summary measure bein meta-analyzed as AUC
  method.tau = "REML", # how the between study variance is estimated
  method.random.ci = "HK" # method for deriving the confidence interval
)
summary(primary_auc)
```

```{r}
# POOLED ESTIMATE ON LOGIT SCALE
library(meta)

primary_logit <- metagen(
  TE               = logit_c, # logit effect estimate gotten from logit_c column
  seTE             = se_logit, # get the logit treatment effect standard error from the Se_logit column
  studlab          = Author, # use the Author column to identify studies
  data             = df_primary_logit, # dataset that contains all the logit transformed data
  method.tau       = "REML",# how the between study variance is estimated
  method.random.ci = "HK", # method for deriving the confidence interval
  prediction       = TRUE # report 95% prediction interval
)
summary(primary_logit)
```

```{r}
# invert pooled value and intervals with inverse-logit
plogis(primary_logit$TE.random)      #perform inverse logit transformation on the pooled estimate
plogis(c(primary_logit$lower.random, primary_logit$upper.random))# perform inverse logit transformation on the confidence interval   
plogis(c(primary_logit$lower.predict, primary_logit$upper.predict)) # perform inverse logit transformation in the prediction interval
```

```{r}
# USING METAMISC LIBRARY AS STATED IN DEBRAY
library(metamisc) # debrey specific package for meta analysis

val <- valmeta( # store result of the velmeta function in the val object
  cstat      = df_primary_logit$Estimate,   # RAW c-statistic provided as logit trransformation is done internally
  cstat.cilb = df_primary_logit$LowerCI,  # RAW lower CI
  cstat.ciub = df_primary_logit$UpperCI, # RAW upper CI
  slab       = df_primary_logit$Index, # use index as the study label
  method     = "REML" # how the between study variance is estimated
)
val
```

```{r}
# FORREST PLOT VABOSE
library(meta)

primary_logit <- metagen(
  TE               = logit_c, # logit effect estimate gotten from logit_c column
  seTE             = se_logit, # get the logit treatment effect standard error from the Se_logit column
  studlab          = Author, # use the Author column to identify studies
  data             = df_primary_logit, # dataset that contains all the logit transformed data
  sm               = "PLOGIT", # tells meta the scale is logit-transformed proportion
  backtransf       = TRUE, # covert logit values to translatable values
  method.tau       = "REML", # how the between study variance is estimated
  method.random.ci = "HK", # method for deriving the confidence interval
  prediction       = TRUE # report 95% prediction interval
)

meta::forest(primary_logit, xlab = "c-statistic") # use forest function in the meta package
```

```{r}
# PLOT FORREST PLOT
meta::forest(primary_logit,
             xlab      = "c-statistic",
             xlim      = c(0.4, 1),
             at        = c(0.5, 0.6, 0.7, 0.8, 0.9, 1.0),
             leftcols  = c("studlab"),
             leftlabs  = c("Study"),
             rightcols = c("effect", "ci", "w.random"),
             rightlabs = c("c-statistic", "95% CI", "Weight"),
             common    = FALSE,
             print.I2  = TRUE,
             print.tau2 = TRUE)
```

## INFLUENCE ANALYSIS

```{r}
#Influenece analysis to be done on logit transformed estimate
# result will be back transformed for translation
# staring with leave-one-out sensitivity analysis
# If droppping a study shifts the pooled c-statistic by >0.05 the result is not stable
library(meta)
loo <- metainf(primary_logit, pooled = "random") # leave one out
loo
meta::forest(loo, backtransf = TRUE, xlab = "c-statistic (leave-one-out)")
```

```{r}
library(metafor)

# attach real study labels so plots use them instead of row numbers
rma_meta <- rma( # random effect meta analysis
  yi = logit_c, #observed effect
  sei = se_logit, # each studies precision and weight
  data = df_primary_logit, # location of variables
         method = "REML", # methos for calculating study estimate
  slab = df_primary_logit$Index)


inf <- influence(rma_meta) # calcualte the influence statistic for every study

print(inf$inf) 

plot(inf) # plot influence diagnostics
# x-axis = each study's contribution to overall heterogeneity
# y-axis = its influence on the pooled estimate
# Build the study key from the model's data, in row order (Baujat labels by row position)
study_key <- paste0(
  seq_len(nrow(df_primary_logit)), # extract the number of rows from the dataframe
  " = ", 
  df_primary_logit$Author) # author name would be combined with row number

plot_settings <- par( # save the graphic setting inside
  mar = c(5, 4, 2, 20), # control lot margins to create better spacing
  xpd = NA) # allow text outside normal plotting region

baujat( # create baugat plot
  rma_meta,
  xlab = "Contribution to overall heterogeneity (Q)",
  ylab = "Influence on pooled c-statistic",
  main = ""
)

usr <- par("usr") # place study key outside the graph for identificarion

text(
  usr[2] + 0.2,
  usr[4],
  labels = paste(
    c("Study key", "", study_key), 
    collapse = "\n"),
  adj = c(0, 1),
  cex = 0.8
)

par(plot_settings)



```

```{r}
# check difference in two dataframe 
library(dplyr)
df_all_only <- anti_join( # keep only rows in first table that do not have a match in the second table
  df_all, # searching table
  df_primary, # reference table
  by = 'Author') # compare the two dataframe using the Author column
print(df_all_only) # print result
```

## SUBGROUP ANALYSIS - Disease group

```{r}
library(meta)

primary_subgroup_disease <- metagen(
  TE               = logit_c, # use logit estimates
  seTE             = se_logit, # use logit SE
  studlab          = Author, # identify by author
  data             = df_primary_logit, # find all in this dataframe
  sm               = "PLOGIT", # inform that the estimate are in logit
  method.tau       = "REML", # how to calculate the between study variance
  method.random.ci = "HK", # how to calcualte the CI 
  subgroup         = Disease_Group, # split the studies according to disease group
  tau.common       = TRUE, # make all studeis in the group share the same tau square by pooling the estimates as few numbers of study would make seperate tau unstable
  prediction       = TRUE # calculate the prediction interval for each group
)

summary(primary_subgroup_disease)
```

FORREST PLOT - SUBGROUP ANALYSIS - Disease group

```{r}
meta::forest(
  primary_subgroup_disease,
  backtransf = TRUE,
  xlab = "c-statistic",
  xlim = c(0.5, 1),
  at = c(0.5, 0.6, 0.7, 0.8, 0.9, 1.0),
  leftcols  = c("studlab", "Sample_size", "Events"),
  leftlabs  = c("Study", "N", "CVD events"),
  rightcols = c("effect", "ci"),
  rightlabs = c("c-statistic", "95% CI"),
  common = FALSE,
  test.subgroup = TRUE,
  print.subgroup.name = FALSE,
  addrows.below.overall = 5,
  colgap = "8mm",
  colgap.left = "6mm",
  colgap.forest.left = "10mm",
  fontsize = 11
)
```

```{r}
# Between-group difference test
primary_subgroup_disease$Q.b.random      # Q between random effect
primary_subgroup_disease$df.Q.b          # degree of freedom
primary_subgroup_disease$pval.Q.b.random # p-value

# Per-subgroup back-transformed estimates
data.frame(
  disease = primary_subgroup_disease$subgroup.levels,
  k       = primary_subgroup_disease$k.w,
  c_stat  = plogis(primary_subgroup_disease$TE.random.w),
  lower   = plogis(primary_subgroup_disease$lower.random.w),
  upper   = plogis(primary_subgroup_disease$upper.random.w),
  I2      = primary_subgroup_disease$I2.w
)
```

## SUBGROUP ANALYSIS BY QRISK VERSION

```{r}
# See version information
table(df_primary_logit$QRISK_Version, useNA = "ifany")
```

```{r}
# subgroup analysis QRISK version
library(meta)

primary_subgroup_version <- metagen(
  TE               = logit_c,
  seTE             = se_logit,
  studlab          = Author,
  data             = df_primary_logit,
  sm               = "PLOGIT",
  method.tau       = "REML",
  method.random.ci = "HK",
  subgroup         = QRISK_Version,      
  tau.common       = TRUE,
  prediction       = TRUE
)

summary(primary_subgroup_version)
```

```{r}
# Extract the between-group test and per-version estimate
# Between-group difference 
primary_subgroup_version$Q.b.random
primary_subgroup_version$df.Q.b
primary_subgroup_version$pval.Q.b.random

# Per-version back-transformed estimates
data.frame(
  version = primary_subgroup_version$subgroup.levels,
  k       = primary_subgroup_version$k.w,
  c_stat  = plogis(primary_subgroup_version$TE.random.w),
  lower   = plogis(primary_subgroup_version$lower.random.w),
  upper   = plogis(primary_subgroup_version$upper.random.w),
  I2      = primary_subgroup_version$I2.w
)
```

```{r}
# widen the bottom margin before plotting
par(mar = c(8, 4, 2, 2))   # bottom, left, top, right — more bottom space

meta::forest(primary_subgroup_version,
             backtransf = TRUE,
             xlab       = "c-statistic",
             xlim       = c(0.5, 1),
             at         = c(0.5, 0.6, 0.7, 0.8, 0.9, 1.0),
             leftcols   = "studlab", leftlabs = "Study",
             rightcols  = c("effect", "ci"), rightlabs = c("c-statistic", "95% CI"),
             common          = FALSE,
             hetstat         = FALSE,   
             test.subgroup   = TRUE,
             overall.hetstat = TRUE,
             fontsize        = 9,          
             spacing         = 1.1)
```

```{r}
# subgroup summary
library(ggplot2)
sg <- data.frame(
  group = primary_subgroup_disease$subgroup.levels,
  k     = primary_subgroup_disease$k.w,
  est   = plogis(primary_subgroup_disease$TE.random.w),
  lo    = plogis(primary_subgroup_disease$lower.random.w),
  hi    = plogis(primary_subgroup_disease$upper.random.w))

ggplot(sg, aes(x = est, y = reorder(group, est))) +
  geom_vline(xintercept = plogis(primary_logit$TE.random),
             linetype = 2, colour = "grey40") +
  geom_errorbarh(aes(xmin = lo, xmax = hi), height = 0.18) +
  geom_point(aes(size = k)) +
  scale_size_continuous(range = c(2, 5), name = "Arms (k)") +
  labs(x = "Pooled c-statistic (95% CI)", y = NULL) +
  theme_minimal(base_size = 11)
```

```{r}
primary_subgroup_disease$Q.b.random
primary_subgroup_disease$df.Q.b
primary_subgroup_disease$pval.Q.b.random
```

## PAIRED QRISK VERSION COMPARISM

```{r}
library(dplyr)
library(tidyr)
library(readr)

# Build the paired set
paired <- df_all %>%
  filter(Index %in% c("19a","19b","26a","26b","29a","29b")) %>%
  mutate(
    Index   = trimws(Index), # trim whitespace from the index
    cohort  = parse_number(Index), # extract the number part from the index
    version = QRISK_Version # store the qrisk version in version
  )

# verify 6 rows, each cohort has one QRISK2 and one QRISK3
print(count(paired, cohort, version))

# Logit-transform each arm c-statistic
paired <- paired %>%
  mutate( # create column for each new logit transformed variable
    logit_c  = log(Estimate / (1 - Estimate)),
    logit_lb = log(LowerCI  / (1 - LowerCI)),
    logit_ub = log(UpperCI  / (1 - UpperCI)),
    se_logit = (logit_ub - logit_lb) / (2 * 1.96)
  )

# Reshape to one row per cohort, QRISK2 and QRISK3 side by side
wide <- paired %>%
  select(cohort, version, Estimate, logit_c, se_logit) %>%
  pivot_wider( # convert to data to wide form
    id_cols     = cohort,
    names_from  = version,
    values_from = c(Estimate, logit_c, se_logit) # variables to be widened
  )

print(names(wide))

# Within-cohort difference (QRISK3 - QRISK2)
wide <- wide %>%
  mutate(
    delta_logit    = logit_c_QRISK3 - logit_c_QRISK2,
    delta_cstat    = Estimate_QRISK3 - Estimate_QRISK2,
    se_delta_indep = sqrt(se_logit_QRISK3^2 + se_logit_QRISK2^2) # SE assuming independence
  )

# Present the paired comparison
wide %>%
  transmute( 
    Cohort                = cohort,
    `QRISK2 c`            = round(Estimate_QRISK2, 3),
    `QRISK3 c`            = round(Estimate_QRISK3, 3),
    `Delta c (3-2)`       = round(delta_cstat, 3),
    `Delta logit (3-2)`   = round(delta_logit, 3),
    `SE (indep, overest)` = round(se_delta_indep, 3)
  )
```

```{r}
# check the average improvement of discrimination between 3 and 2
mean_delta_logit <- mean(wide$delta_logit)
cat("Mean within-cohort logit difference (QRISK3 - QRISK2):",
    round(mean_delta_logit, 4), "\n")
```

## SUBGROUP ANALYSIS (PROBAST)

```{r}
# Which column holds the overall PROBAST ROB verdict, and how is it coded?
table(df_primary_logit$Overall, useNA = "ifany")
```

```{r}
# ROB SUBGROUP ANALYSIS
library(meta)

primary_subgroup_rob <- metagen(
  TE               = logit_c,
  seTE             = se_logit,
  studlab          = Author,
  data             = df_primary_logit,
  sm               = "PLOGIT",
  method.tau       = "REML",
  method.random.ci = "HK",
  subgroup         = Overall,    
  tau.common       = TRUE,
  prediction       = TRUE
)

summary(primary_subgroup_rob)
```

```{r}
# Between-group test and per-group estimates
# Between-group difference
primary_subgroup_rob$Q.b.random
primary_subgroup_rob$df.Q.b
primary_subgroup_rob$pval.Q.b.random

# Per-ROB-group back-transformed estimates
data.frame(
  rob    = primary_subgroup_rob$subgroup.levels,
  k      = primary_subgroup_rob$k.w,
  c_stat = plogis(primary_subgroup_rob$TE.random.w),
  lower  = plogis(primary_subgroup_rob$lower.random.w),
  upper  = plogis(primary_subgroup_rob$upper.random.w),
  I2     = primary_subgroup_rob$I2.w
)
```

```{r}
# FORST PLOT FOR ROB
par(mar = c(8, 4, 2, 2))

meta::forest(
  primary_subgroup_rob,
  backtransf = TRUE,
  xlab = "c-statistic",
  xlim = c(0.5, 1),
  at = c(0.5, 0.6, 0.7, 0.8, 0.9, 1.0),

  leftcols = "studlab",
  leftlabs = "Study",

  rightcols = c("effect", "ci"),
  rightlabs = c("c-statistic", "95% CI"),

  common = FALSE,

  hetstat = FALSE,
  overall.hetstat = TRUE,
  test.subgroup = TRUE,

  fontsize = 9,
  spacing = 1.1,

  addrows.below.overall = 2
)
```

## SENSITIVITY ANALYSIS DROPING LOW STUDIES

```{r}
df_lowrob <- df_primary_logit[df_primary_logit$Overall == "LOW", ]

primary_lowrob <- metagen(
  TE = logit_c, 
  seTE = se_logit, 
  studlab = Index, 
  data = df_lowrob,
  sm = "PLOGIT", 
  method.tau = "REML", 
  method.random.ci = "HK", 
  prediction = TRUE
)
summary(primary_lowrob)
```

## SENSITIVITY ANALYSIS FOR DISCRIMINATION METRIC TYPE

```{r}
# CHECK TYPE
table(df_primary_logit$Dscrimination, useNA = "ifany")

```

```{r}
# LEAVE OUT ONLY HARRELLS 
library(meta)

# Keep only the AUC-metric studies (drop the three Harrell's C arms)
df_auc_only <- df_primary_logit[df_primary_logit$Dscrimination == "AUC", ]
nrow(df_auc_only) 

sens_auc_only <- metagen(
  TE               = logit_c,
  seTE             = se_logit,
  studlab          = Index,
  data             = df_auc_only,
  sm               = "PLOGIT",
  method.tau       = "REML",
  method.random.ci = "HK",
  prediction       = TRUE
)
summary(sens_auc_only)
```

## SUBGROUP ANALYSIS UK vs Non UK

```{r}
table(df_primary_logit$Country, useNA = "ifany")

```

```{r}
library(dplyr)
library(meta)

# Collapse to UK vs non-UK
df_primary_logit <- df_primary_logit %>%
  mutate(UK_group = ifelse(Country == "UK", "UK", "Non-UK"))

table(df_primary_logit$UK_group)

primary_subgroup_uk <- metagen(
  TE               = logit_c,
  seTE             = se_logit,
  studlab          = Index,
  data             = df_primary_logit,
  sm               = "PLOGIT",
  method.tau       = "REML",
  method.random.ci = "HK",
  subgroup         = UK_group,
  tau.common       = TRUE,
  prediction       = TRUE
)
summary(primary_subgroup_uk)
```

## SUGGROUP ANALYSIS - FOLLOW UP

```{r}
# See the follow-up values across the pool
df_primary_logit %>%
  select(Index, Disease_Group, `Follow-Up`, Estimate) %>%
  arrange(`Follow-Up`) %>%
  print(n = Inf)
```

```{r}
library(dplyr)
library(meta)

fu_under10 <- c("1", "36", "3", "16", "40", "23", "39")

df_primary_logit <- df_primary_logit %>%
  mutate(Follow_group = ifelse(Index %in% fu_under10, "Under 10y", "10y or more"))

table(df_primary_logit$Follow_group, useNA = "ifany")

sens_followup <- metagen(
  TE = logit_c, 
  seTE = se_logit, 
  studlab = Index, 
  data = df_primary_logit,
  sm = "PLOGIT", 
  method.tau = "REML", 
  method.random.ci = "HK",
  subgroup = Follow_group, 
  tau.common = TRUE, 
  prediction = TRUE
)
summary(sens_followup)
```

## DESCRIPTIVE ANALYSIS - HELLERS K

```{r}
library(dplyr)

carma <- df_all %>%
  filter(Index %in% c("5a", "5b", "5c")) %>%
  select(Index, Disease_Group, Country, Sample_size,
         Estimate, LowerCI, UpperCI, Dscrimination, `Follow-Up`)

carma
```

## SENSITIVITY ANALYSIS - CROSS-SECTIONAL

```{r}
library(dplyr)

# Pulling the candidate cross-sectional surrogate-outcome studies from df_all
cross_sec <- df_all %>%
  filter(Outcome_type != "Appropriate") %>%   
  select(Index, Disease_Group, Outcome_type, Outcome_Description,
         `Follow-Up`, Estimate, LowerCI, UpperCI, Dscrimination)

cross_sec
```

```{r}
library(dplyr)

cross_sec <- df_all %>%
  filter(Index %in% c("14","18","21","22","28","32","37","27b")) %>% 
  select(Index, Disease_Group, Outcome_Description, Dscrimination,
         Estimate, LowerCI, UpperCI, `Follow-Up`, Confidence_interval)

cross_sec
```

```{r}
# Discrimination for subclinical carotid atherosclerosis
library(dplyr)
library(meta)

plaque <- df_all %>%
  filter(Index %in% c("14","18","22","28","37","27b")) %>%   # carotid plaque only; 21 & 32 excluded
  mutate(
    logit_c  = log(Estimate / (1 - Estimate)),
    logit_lb = log(LowerCI  / (1 - LowerCI)),
    logit_ub = log(UpperCI  / (1 - UpperCI)),
    se_logit = (logit_ub - logit_lb) / (2 * 1.96)
  )

nrow(plaque) 

pool_plaque <- metagen(
  TE = logit_c, 
  seTE = se_logit, 
  studlab = Index, 
  data = plaque,
  sm = "PLOGIT", 
  method.tau = "REML", 
  method.random.ci = "HK", 
  prediction = TRUE
)
summary(pool_plaque)
```

```{r}
df_primary_logit %>%
  select(Index, Sample_size, Study_Year, `Follow-Up`
        ) %>%
  print(n = Inf)
```

## SUBGROUP ANALYSIS - MULTI-VS-SINGLE

```{r}
library(meta)

# First check how the groups split and that the labels are clean
table(df_primary_logit$`Single_vs_multi-centre`, useNA = "ifany")
```

```{r}
primary_subgroup_centre <- metagen(
  TE               = logit_c,
  seTE             = se_logit,
  studlab          = Index,
  data             = df_primary_logit,
  sm               = "PLOGIT",
  method.tau       = "REML",
  method.random.ci = "HK",
  subgroup         = `Single_vs_multi-centre`,
  tau.common       = TRUE,
  prediction       = TRUE
)
summary(primary_subgroup_centre)
```

```{r}
data.frame(
  Q  = primary_subgroup_centre$Q.b.random,
  df = primary_subgroup_centre$df.Q.b,
  p  = primary_subgroup_centre$pval.Q.b.random
)

data.frame(
  centre = primary_subgroup_centre$subgroup.levels,
  k      = primary_subgroup_centre$k.w,
  c_stat = plogis(primary_subgroup_centre$TE.random.w),
  lower  = plogis(primary_subgroup_centre$lower.random.w),
  upper  = plogis(primary_subgroup_centre$upper.random.w),
  I2     = primary_subgroup_centre$I2.w
)


```

# METAREGRESSION

## Event meta regression

```{r}
library(metafor)

# Log-transform events
df_primary_logit$log_events <- log(df_primary_logit$Events)

# Refit the primary model in metafor as the reference
m_base <- rma(
  yi = logit_c, 
  sei = se_logit, 
  data = df_primary_logit, 
  method = "REML")
m_base
# Univariable meta-regression on log(events)
mr_events <- rma(
  yi = logit_c, 
  sei = se_logit,
  mods = ~ log_events,
  data = df_primary_logit, 
  method = "REML")

mr_events
```

```{r}
mr_events
```

### To make the slope interpretation on the c-statistic scale

```{r}
# Predicted c-statistic across the range of events
# use the fitted metaregression model to estimate the c-statistic for studies at the event number
newpts <- data.frame(log_events = log(c(20, 100, 500, 5000, 25000))) # create dataframe for hypothetical moderators/event number
prediction <- predict( # function to calculate predicted values from fitted model
  mr_events, # fitted regression model
  newmods = as.matrix(newpts$log_events), # convert moderator to a matrix to feed to function
  transf = plogis
  )

# build result table
data.frame(events = c(20, 100, 500, 5000, 25000),
           c_stat = round(prediction$pred, 3),
           lower  = round(prediction$ci.lb, 3),
           upper  = round(prediction$ci.ub, 3))
```

### **Bubble plot** for the figure

```{r}
regplot(mr_events, 
        mod = "log_events", # name of relevant data in the model object
        xlab = "log(number of events)", 
        ylab = "logit c-statistic",
        las = 1, # axis label
        digits = 2, # number of decimal places in the plot
        bg = "orange")

```

## Event rate - meta regression

```{r}
df_primary_logit$event_rate <- df_primary_logit$Events / df_primary_logit$Sample_size

mr_rate <- rma(
  yi = logit_c, 
  sei = se_logit,
  mods = ~ event_rate,
  data = df_primary_logit, 
  method = "REML")

mr_rate
```

## Sample Size - meta regression

```{r}
df_primary_logit$log_n <- log(df_primary_logit$Sample_size)

mr_sample_size <- rma(
  yi = logit_c, 
  sei = se_logit,
  mods = ~ log_n,
  data = df_primary_logit, 
  method = "REML")

mr_sample_size
```

```{r}
library(meta)

# Build the numbered key from the pooled object (row order = plot order)
key <- data.frame(
  No    = seq_along(primary_logit$studlab), # create number sequence
  Study = primary_logit$studlab,
  c_stat = round(plogis(primary_logit$TE), 3) # study effect column
)

# Funnel plot labelled with numbers
meta::funnel(primary_logit,
             studlab     = as.character(key$No),# display numbers
             cex.studlab = 0.7, # font size
             pos.studlab = 4, # label to the right of each point
             xlab        = "logit c-statistic",
             backtransf  = FALSE,
             col = "orange"
)
# Print the mapping to place in the figure caption or legend
print(key, row.names = FALSE)

# Egger-type test for funnel asymmetry
metabias(primary_logit, method.bias = "linreg", k.min = 10)
paste(paste0(key$No, " = ", key$Study), collapse = "; ") # caption for keys
```

```{r}
#traffic light plot
library(dplyr); library(tidyr); library(ggplot2)

rob <- df_primary_logit %>%
  select(Index, Author, Participants, Predictors2, Outcome3, Analysis, Overall) %>%
  pivot_longer(c(-Index, -Author), names_to = "Domain", values_to = "Rating") %>%
  mutate(
    Domain = recode(Domain, Predictors2 = "Predictors", Outcome3 = "Outcome"),
    Domain = factor(Domain,
             levels = c("Participants","Predictors","Outcome","Analysis","Overall")),
    Rating = factor(toupper(trimws(Rating)), levels = c("LOW","UNCLEAR","HIGH")))

ggplot(rob, aes(x = Domain, y = factor(Author, levels = rev(unique(Author))))) +
  geom_tile(aes(fill = Rating), colour = "white", linewidth = 0.8) +
  geom_text(aes(label = recode(as.character(Rating),
                LOW = "+", UNCLEAR = "?", HIGH = "\u2212")), size = 3.5) +
  scale_fill_manual(values = c(LOW = "#4DAF4A", UNCLEAR = "#FFD92F", HIGH = "#E41A1C")) +
  labs(x = NULL, y = "Study", fill = "Risk of bias") +
  theme_minimal(base_size = 10) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

```{r}
png("metareg_panel.png", width = 2400, height = 900, res = 200)
par(mfrow = c(1, 3), mar = c(5, 5, 3, 2))
regplot(mr_events, mod = "log_events", xlab = "log(events)",
        ylab = "logit c-statistic", main = "A", bg = "grey")
regplot(mr_sample_size, mod = "log_n", xlab = "log(sample size)",
        ylab = "", main = "B", bg = "grey")
regplot(mr_rate, mod = "event_rate", xlab = "Event rate",
        ylab = "", main = "C", bg = "grey")
dev.off()
```

```{r}
pkgs <- c("meta", "metafor", "metamisc", "readxl", "writexl",
          "dplyr", "stringr", "tidyr", "ggplot2", "gt")

purpose <- c(
  meta     = "primary pooling (metagen), subgroups, leave-one-out, funnel plot",
  metafor  = "case-deletion diagnostics, Baujat plot, meta-regression (rma)",
  metamisc = "reference implementation of the Debray framework (valmeta)",
  readxl   = "data import",
  writexl  = "data export",
  dplyr    = "data preparation",
  stringr  = "text parsing of the AUC field",
  tidyr    = "data reshaping",
  ggplot2  = "risk-of-bias figures",
  gt       = "formatted tables"
)

env <- data.frame(
  Package = pkgs,
  Version = vapply(pkgs, function(p) as.character(packageVersion(p)), character(1)),
  Purpose = purpose[pkgs],
  row.names = NULL
)

cat("R version:", paste0(R.version$major, ".", R.version$minor), "\n")
cat("Platform :", R.version$platform, "\n\n")
print(env, right = FALSE)
```

```{r}
sessionInfo()
```

```{r}
citation()            # R itself
citation("meta")
citation("metafor")
citation("metamisc")
```
